---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

pkgload::load_all(".", quiet = TRUE, export_all = FALSE, helpers = FALSE)
```

# sigProCon

<!-- badges: start -->
<!-- badges: end -->

It's common to have a signature of genes, then to give each sample in a bulk rna-seq dataset a score reflecting the activity of that signature, such as by GSVA.

The goal of sigProCon is to identify which genes from a signature are contributing most to the sample signature activity scores across all samples.  

For example, one may have a signature of genes that are upregulated during the process of epithelial to mesencymal transition (EMT), which is associated with tumor metastasis.
One can use a tool (such as GSVA) to assign a score to each patient in the TCGA dataset based on each patient's gene expression profile, that corresponds to how 'active' the EMT signature is in each patient.  It would suggest that a patient with a higher score has more activity of the genes in the EMT geneset, and thus may have a more aggressive tumor.  Some genes will be very correlated with the score, suggesting their larger role in EMT in this particular set of samples, while some genes in the EMT signature may have very little correlation.

This package simply takes the correlation of each gene in the signature with the signature score, to identify the genes that are contributing most to the score. It also supports the visualization of the results as an annotated heatmap and a KS plot.

## Installation

You can install sigProCon from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("montilab/sigProCon")
```

## Example with toy sample data:

sigProCon ships with a small toy signature and toy expression set for quick experimentation; here we check their sizes:

```{r toy-example, message=FALSE, warning=FALSE}
data(toy_signature)
data(toy_eset)
print(length(toy_signature))
print(dim(toy_eset))
```

### View toy data tables:

`signature_projection_contributors()` returns, among other things, each gene's correlation to the signature score (`score_cor`) and each sample's computed score (`sig_score`):

```{r toy-tables, message=FALSE, warning=FALSE}
toy_output <- sigProCon::signature_projection_contributors(
  eset = toy_eset,
  signature = list(signature1 = toy_signature)
)
## Show some of the genes' correlation to the signature score
print(head(toy_output$score_cor))

## Show some of the signature scores
print(head(toy_output$sig_score))
```

### View toy heatmap that plots all signature and non-signature genes:

`spc_heatmap_all()` draws every gene in the dataset ranked by its correlation to the signature score, with signature genes and KS leading-edge hits flagged on the right:

```{r toy-full.heatmap, dpi=300}
## plot full heatmap
sigProCon::spc_heatmap_all(
  eset = toy_eset,
  spc_out = toy_output
)
```

### View toy heatmap that plots only signature genes:

`spc_heatmap_sig()` shows the same information restricted to just the signature genes, which is easier to read when the full gene set is large:

```{r toy-sig.heatmap, dpi=300}
## plot signature heatmap
sigProCon::spc_heatmap_sig(
  eset = toy_eset,
  spc_out = toy_output
)
```

### View toy KS plot of signature genes and scores:

The KS plot shows where the signature genes fall along the full ranked gene list (by correlation to the score) — a peak early in the list means the signature genes are concentrated among the most correlated genes, similar to a GSEA enrichment plot:

```{r toy-ksplot, dpi=300}
print(toy_output$ks$plot)
```

Under the hood, the signature score is computed by `omics_signature_score()`, which supports three methods: `"GSVA"` (the default), `"eigengene"` (the module's first eigengene via WGCNA), and `"pc"` (a direct principal-component score).

#### Passing a pre-computed score

If you already have a signature score (or want to reuse one across multiple calls), pass it directly via `sig_score` to skip recomputing it:

```{r toy-tables-precomputed, message=FALSE, warning=FALSE}
toy_score <- sigProCon::omics_signature_score(
  eset = toy_eset,
  signature = list(signature1 = toy_signature),
  method = "GSVA"
)
toy_score_eigengene <- sigProCon::omics_signature_score(
  eset = toy_eset,
  signature = list(signature1 = toy_signature),
  method = "eigengene"
)
## call w/ pre-computed score 
toy_output_precomp <- sigProCon::signature_projection_contributors(
  eset = toy_eset,
  signature = list(signature1 = toy_signature),
  sig_score = toy_score
)
## same results
isTRUE(all.equal(toy_output$score_cor, toy_output_precomp$score_cor))
isTRUE(all.equal(toy_output$sig_score, toy_output_precomp$sig_score))
```

## Example with package sample data (`eset` and `signature`):

The package also ships a larger, more realistic dataset (`eset` and `signature`) for demonstration:

```{r plots, message=FALSE, warning=FALSE, results='hide' }
#with our sample data
data(sigProCon::signature)
data(sigProCon::eset)

output <- sigProCon::signature_projection_contributors(
  eset = sigProCon::eset, 
  signature = sigProCon::signature)
```

### View tables:

As with the toy example, `score_cor` and `sig_score` hold the per-gene correlations and per-sample scores:

```{r tables}
## Show some of the genes' correlation to the signature score
print(head(output$score_cor))
## Show some of the signature scores
print(head(output$sig_score))
```

### View heatmap that plots only signature genes:

Signature-only view for the full dataset:

```{r sig.heatmap, dpi=300}
## plot signature heatmap
sigProCon::spc_heatmap_sig(
  eset = sigProCon::eset,
  spc_out = output
)
```

### View heatmap that plots all signature and non-signature genes:

All-genes view for the full dataset:

```{r full.heatmap, dpi=300}
## plot full heatmap
sigProCon::spc_heatmap_all(
  eset = sigProCon::eset,
  spc_out = output
)
```

The large number of genes in the dataset makes it difficult to clearly see the signature hits (_leadedge_ on the right). We can subsample the number of genes to be displayed (to be the union of the top `subsample` genes by MAD and the signature genes). We can also display the top correlated genes (using `leadedge_label_n`).

```{r full.heatmap.subsample, dpi=300}
## plot full heatmap
sigProCon::spc_heatmap_all(
  eset = sigProCon::eset,
  spc_out = output,
  subsample = 3000,
  leadedge_label_n = 5
)
```

### View KS plot of signature genes and scores:

KS plot for the full dataset:

```{r ksplot, dpi=300}
print(output$ks$plot)
```
